Developing a preprint culture in biology

### 2012-10-18 ms.tex
final final editing
Add ref to Haldane's sieve

Open preprint servers offer a great opportunity for open science, especially if
the community embraces the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work \cite{lom12}. These
initiatives are important to fully exploit the potential of open preprint
servers. Posting preprints online increases the community of available informal
peer reviewers, and uses the internet for its original community-building
purposes.

@@

The ongoing discussions on the publication process, peer-reviewing and
alternative publication models are all symptoms of the current uneasiness of
scientists with the ever growing obsession with bibliographic metrics such as
the impact factor \cite{Fisher2012}. There is pressure on researchers to orient
their publication strategy in order to maximize their number of publications and
total citations. A well-known consequence is to submit manuscripts first to the
most prestigious journals, and then resubmit to ``lower level'' journals as they
are rejected. The numerous negative impacts of such behavior have been discussed
in depth \cite{hoc09} and include a long delay between the time a manuscript is
finished to its publication.  They all contribute somehow in a general slowing
down of scientific progress.  Research activities and the publication process
are drifting away from their fundamental object, namely the diffusion of novel
scientific discoveries.

Developing a preprint culture in biology will not solve all problems with the
current publication process. However, It might significantly reduce its negative
consequences. The role of peer-reviewing is to judge the scientific quality of a
study. It is the first barrier against the fraudulent and poor quality science
susceptible to impede scientific progress. In practice, the peer-review system
is not only used to evaluate scientific quality but also to judge pertinence. On
the other hand, preprints are not filtered, neither for their quality nor their
pertinence. Widespread adoption of preprint servers has the potential to shift
the diffusion strategy: journals would remain important to validate
publications, but the relevance of a study, its contribution to science, should
only be judged by many more readers than the typical two-four anonymous
reviewers. With a shift in the diffusion strategy, the role of traditional
journals and their editors would be to showcase scientific discoveries for
specialized readership.

>>>>

Open preprint servers offer a great opportunity for open science, especially if
the community embraces the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}) \cite{coo12}, a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work \cite{lom12}. These
initiatives are important to fully exploit the potential of open preprint
servers. Posting preprints online increases the community of available informal
peer reviewers, and uses the internet for its original community-building
purposes.

@@

The ongoing discussions on the publication process, peer-reviewing and
alternative publication models are all symptoms of the current uneasiness of
scientists with the ever growing obsession with bibliographic metrics such as
the impact factor \cite{Fisher2012}. There is pressure on researchers to orient
their publication strategy in order to maximize their number of publications and
total citations. A well-known consequence is to submit manuscripts first to the
most prestigious journals, and then resubmit to ``lower level'' journals as they
are rejected. The numerous negative impacts of such behavior have been discussed
in depth \cite{hoc09} and include a long delay between the time a manuscript is
finished to its publication.  They all contribute in a general slowing down of
scientific progress.  Research activities and the publication process are
drifting away from their fundamental object, namely the diffusion of novel
scientific discoveries.

Developing a preprint culture in biology will not solve all problems with the
current publication process. However, it might significantly reduce its negative
consequences. The role of peer-reviewing is to judge the scientific quality of a
study. It is the first barrier against the fraudulent and poor quality science
susceptible to impede scientific progress. In practice, the peer-review system
is not only used to evaluate scientific quality but also to judge pertinence. On
the other hand, preprints are not filtered, neither for their quality nor their
pertinence. Widespread adoption of preprint servers has the potential to shift
the diffusion strategy: journals would remain important to validate
publications, but the relevance of a study, its contribution to science, should
only be judged by many more readers than the typical two-four anonymous
reviewers. With a shift in the diffusion strategy, the role of traditional
journals and their editors would be to showcase scientific discoveries for
specialized readership.
